Supplementary material S1: Leaf lifespan model

### 2018-12-13 A01-mortality.Rmd
Draft 0.1.9010 Co-authors sharing

---
title: "Nino mortality issue"
date: '`r Sys.Date()`'
output:
  bookdown::pdf_document2:
    number_sections: false
    toc: false
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

>>>>

---
title: "A01 - Nino mortality issue"
date: '`r Sys.Date()`'
output:
  bookdown::pdf_document2:
    number_sections: false
    toc: false
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

### 2018-12-13 A02-Resilience.Rmd
Draft 0.1.9010 Co-authors sharing

---
title: "Supplementary material S2: Resilience indice"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
  bookdown::word_document2: default
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

@@

This document introduces and explains the recovery function $R(X)$ and the resilience indice $I_R$ definitions (figure \@ref(fig:IrGraphFinal)). Resilience concept has met numerous definitions, but resilience in ecology can encompass both ecosystem ability to adsorb and recover from disturbance, resistance and recovery respectively [@Gunderson2010].

```{r IrGraphFinal, fig.cap="Definition of recovery function and resilience indice shown on aboveground biomass dynamics of a control and disturbed ecosystem before and after the disturbance event."}
knitr::include_graphics("images/final.png")
```

@@

When $R(t)$ approaches $R_{eq}=1$, the disturbed and control systems are considered equivalent, indicating a perfect recovery. We then calculated the euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$. Because the intensity of disturbance was the same for all simulations, our analysis did not include any ecosystem resistance, i.e. the ecosystem capacity to absorb a disturbance event,. Consequently **in this particular case**, we can use the cummulated recovery as an indice of ecosystem resilience, further defined as $I_R$:

@@

```{r datatrans, echo=FALSE, fig.cap='Ecosystem outputs data transformation. Ecosystem outputs (**A**) are normalized by the control value over time to calculate recovery (**B**); recovery of different ecosystem outputs is then used in a multidimensional space to caluclate ecosystem distance to equilibrium (**C**); finally distance at equilibrium is integrated over time in a cumulative sum (**D**). ', cache=TRUE, fig.height=8, fig.width=12, message=FALSE}
cols <- RColorBrewer::brewer.pal(3, "YlOrRd")
g_agb <- ecosystem %>% 
  left_join(., doe, by = 'sim') %>%
  filter(SR == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = agb/1000, color = disturbance)) +
  geom_point() +
  xlab('Time (years)') +
  ylab('Aboveground biomass (tonC/ha)') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                       values=c("#999999", cols))
g_resilience <- resilience %>% 
  left_join(., doe, by = 'sim') %>%
  filter(SR == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = agb, color = disturbance)) +
  geom_hline(yintercept = 1, col = 'red') +
  geom_point() +
  xlab('Time (years)') +
  ylab('Resilience of aboveground biomass') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols)
g_deq <- Deq %>% 
  left_join(., doe, by = 'sim') %>%
  filter(SR == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = structure, 
             color = disturbance, fill = disturbance)) +
  geom_hline(yintercept = 0, col = 'red') +
  geom_point() +
  geom_area(alpha = 0.2) +
  xlab('Time (years)') +
  ylab('Distance to forest structure equilibrium\n(AGB, BA, N, N10 and N30)') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols) +
  scale_fill_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols)
g_Ieq <- Deq %>% 
  group_by(sim) %>% 
  arrange(time) %>% 
  mutate_at(vars(-time), cumsum) %>% 
  left_join(., doe, by = 'sim') %>%
  filter(SR == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = structure, color = disturbance)) +
  geom_point() +
  xlab('Time (years)') +
  ylab('Cummulative integral of\ndistance to forest structure equilibrium\n(AGB, BA, N, N10 and N30)') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols)
cowplot::plot_grid(g_agb, g_resilience, g_deq, g_Ieq, labels = LETTERS[1:4])
```

>>>>

---
title: "A02 - Resilience index"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
  bookdown::word_document2: default
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

@@

This document introduces and explains the recovery function $R(X)$ and the resilience index $I_R$ definitions (figure \@ref(fig:IrGraphFinal)). Resilience concept has met numerous definitions, but resilience in ecology can encompass both ecosystem ability to adsorb and recover from disturbance, resistance and recovery respectively [@Gunderson2010].

```{r IrGraphFinal, fig.cap="Definition of recovery function and resilience indice shown on aboveground biomass dynamics of a control and disturbed ecosystem before and after the disturbance event."}
knitr::include_graphics("images/figure3.png")
```

@@

When $R(t)$ approaches $R_{eq}=1$, the disturbed and control systems are considered equivalent, indicating a perfect recovery. We then calculated the euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$. Because the intensity of disturbance was the same for all simulations, our analysis did not include any ecosystem resistance, i.e. the ecosystem capacity to absorb a disturbance event,. Consequently **in this particular case**, we can use the cummulated recovery as an index of ecosystem resilience, further defined as $I_R$:

@@

```{r datatrans, echo=FALSE, fig.cap='Ecosystem outputs data transformation. Ecosystem outputs (**A**) are normalized by the control value over time to calculate recovery (**B**); recovery of different ecosystem outputs is then used in a multidimensional space to caluclate ecosystem distance to equilibrium (**C**); finally distance at equilibrium is integrated over time in a cumulative sum (**D**). ', cache=TRUE, fig.height=8, fig.width=12, message=FALSE}
cols <- RColorBrewer::brewer.pal(3, "YlOrRd")
g_agb <- ecosystem %>% 
  ungroup() %>% 
  left_join(., doe, by = 'sim') %>%
  filter(SR == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = agb/1000, color = disturbance)) +
  geom_point() +
  xlab('Time (years)') +
  ylab('Aboveground biomass (tonC/ha)') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                       values=c("#999999", cols))
g_resilience <- resilience %>% 
  ungroup() %>%
  left_join(., doe, by = 'sim') %>%
  filter(SR == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = agb, color = disturbance)) +
  geom_hline(yintercept = 1, col = 'red') +
  geom_point() +
  xlab('Time (years)') +
  ylab('Resilience of aboveground biomass') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols)
g_deq <- Deq %>% 
  left_join(., doe, by = 'sim') %>%
  ungroup() %>%
  filter(SR == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = structure, 
             color = disturbance, fill = disturbance)) +
  geom_hline(yintercept = 0, col = 'red') +
  geom_point() +
  geom_area(alpha = 0.2) +
  xlab('Time (years)') +
  ylab('Distance to forest structure equilibrium\n(AGB, BA, N, N10 and N30)') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols) +
  scale_fill_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols)
g_Ieq <- Deq %>% 
  group_by(sim) %>% 
  arrange(time) %>% 
  mutate_at(vars(-time), cumsum) %>% 
  left_join(., doe, by = 'sim') %>%
  filter(SR == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = structure, color = disturbance)) +
  geom_point() +
  xlab('Time (years)') +
  ylab('Cummulative integral of\ndistance to forest structure equilibrium\n(AGB, BA, N, N10 and N30)') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols)
cowplot::plot_grid(g_agb, g_resilience, g_deq, g_Ieq, labels = LETTERS[1:4])
```

### 2018-12-13 A03-RNE.Rmd
Draft 0.1.9010 Co-authors sharing

---
title: "Supplementary material S3: Biodiversity net effect and resilience"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: true
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

>>>>

---
title: "A03 - Biodiversity net effect and resilience"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: true
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

### 2018-12-13 A04-LM.Rmd
Draft 0.1.9010 Co-authors sharing

---
title: "Supplementary material S4: Linear models"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
  bookdown::word_document2: default
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

>>>>

---
title: "A04: Linear models"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
  bookdown::word_document2: default
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

### 2018-12-13 A05-Biodiversity.Rmd
Draft 0.1.9010 Co-authors sharing

---
title: "Supplementary material S5: Biodiversity effect results"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
  bookdown::word_document2: default
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

@@

This document explores graphically the result of the biodiversity net effect partitionning. More especially we want to look if the representation from figure \@ref(fig:gBE) is correct, with the dynamic biodiversity net effect normalized by the control across ecosystems. In order to do so, we will start by looking into the dynamics of a single ecosystem \@ref(fig:gBEsingle), we will then move to all ecosystems not normalized by the control \@ref(fig:gBEunorm), before validating our current representation \@ref(fig:gBE).

@@

The net effect of biodiversity on ecosystem recovery was partitioned between selection and complementarity effect. We are thus able for one community to look at aboveground biomass dynamic after and before disturbance in control and disturbed ecosystems and correspondig net biodiversity effect, selection effect and complementarity effect (figure \@ref(fig:gBEsingle)). Similarly to the aboveground biomass of the ecosystem, the net biodiversity effect and selection effect decrease following the disturbance, whereas the complementarity effect directly increases following the disturbance. Interestingly, we can see that this pattern of decrease of selection effect after disturbance and recovery after is similar to the effect of succession in forest early stages, when the new cohort arrives selection effect suddenly decreases and complementarity effect insure a small net effect (first century, sub-figure on left-bottom on figure \@ref(fig:gBEsingle)). Anyway the pattern is stronger after a disturbance, maybe due to an increased diversity allowing a stronger diversity effect.

@@

If we aggregate the result of all community with 125 species, we still observe the same pattern for all simulations compard to the single previous one (figure \@ref(fig:gBEunorm)). But due to an increase of biodiversity net effect through recovery time, complementarity importance in early recovery is less visible over a long period. Consequently normalizing by control biodiversity net effect can help avoiding this issue.

@@

The net effect of biodiversity on ecosystem recovery was partitioned between selection and complementarity effect after normalizing using undisturbed control simulation ($disturb_{intensity}=0\%$, figure \@ref(fig:gBE)). Thanks to previous graphical exploration, we can affirm that normalizing by the control is not distorting the result. And this last figure brings the clearer message regarding complementarity and selection effect role in the resilience of the biodiveristy net effect.

>>>>

---
title: "A05: Biodiversity effect results"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
  bookdown::word_document2: default
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

@@

This document explores graphically the result of the biodiversity net effect partitioning. More especially we want to look if the representation from figure \@ref(fig:gBE) is correct, with the dynamic biodiversity net effect normalized by the control across ecosystems. In order to do so, we will start by looking into the dynamics of a single ecosystem \@ref(fig:gBEsingle), we will then move to all ecosystems not normalized by the control \@ref(fig:gBEunorm), before validating our current representation \@ref(fig:gBE).

@@

The net effect of biodiversity on ecosystem recovery was partitioned between selection and complementarity effect. We are thus able for one community to look at aboveground biomass dynamic after and before disturbance in control and disturbed ecosystems and corresponding net biodiversity effect, selection effect and complementarity effect (figure \@ref(fig:gBEsingle)). Similarly to the aboveground biomass of the ecosystem, the net biodiversity effect and selection effect decrease following the disturbance, whereas the complementarity effect directly increases following the disturbance. Interestingly, we can see that this pattern of decrease of selection effect after disturbance and recovery after is similar to the effect of succession in forest early stages, when the new cohort arrives selection effect suddenly decreases and complementarity effect insure a small net effect (first century, sub-figure on left-bottom on figure \@ref(fig:gBEsingle)). Anyway the pattern is stronger after a disturbance, maybe due to an increased diversity allowing a stronger diversity effect.

@@

If we aggregate the result of all community with 125 species, we still observe the same pattern for all simulations compared to the single previous one (figure \@ref(fig:gBEunorm)). But due to an increase of biodiversity net effect through recovery time, complementarity importance in early recovery is less visible over a long period. Consequently normalizing by control biodiversity net effect can help avoiding this issue.

@@

The net effect of biodiversity on ecosystem recovery was partitioned between selection and complementarity effect after normalizing using undisturbed control simulation ($disturb_{intensity}=0\%$, figure \@ref(fig:gBE)). Thanks to previous graphical exploration, we can affirm that normalizing by the control is not distorting the result. And this last figure brings the clearer message regarding complementarity and selection effect role in the resilience of the biodiversity net effect.

### 2018-12-13 S1-LL.Rmd
Draft 0.1.9010 Co-authors sharing

---
title: "Supplementary material S1: Leaf lifespan model"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

@@

We compiled functional traits from GLOPNET [@wright_worldwide_2004], TRY [@Kattge2011], and DRYAD [@chave_towards_2009] databases (see \@ref(tab:A2traits)). We kept dataset given by GLOPNET as origin dataset for observations. Dataset defined as origin corresponded to leaf lifespan ($LL$) and most of the time to leaf mass per area ($LMA$) and leaf nitrogen content per leaf dry mass ($Nmass$). We measured variable importance in functionnal traits to explain leaf lifespan with an out-of the bag method applied on a random forest. Then, we used a bayesian apporach to test different models with growing level of complexity. We retained the model with the best tradeoff between model complexity (number of parameters $K$), convergence, likelihood, and prediction quality (root mean sqaure error of prediciton $RMSEP$). We finally tested the new allometry obtained with the selected model with TROLL simulations.

@@

Leaf lifespan $LL$ follows a lognormal law with location infered from leaf lifespan $LMA$, nitrogen content $N$ and wood specific gravity $wsg$ and a scale $\sigma$. Each ${\beta_i}_d$ is following a normal law located on $\beta_i$ with a scale of $\sigma_i$. All $\beta_i$, $\sigma_i$, and $\sigma$ are assumed without presemption following a gamma law. $d$ represents the dataset random effects and encompass environmental and protocol variations.

@@

Simulations are validating that this new allometry resolve the issue of unrealistically low aboveground biomass for low LMA species due to an early death of individuals inside simulations. For instance with this allometry Symphonia sp 1 (a low LMA species) is now reaching a realistic aboveground biomass above 400 $tonC.ha^{-1}$ and realistic diameter and age distribution inside the final population.

>>>>

---
title: "Supplementary material S1: Leaf lifespan model"
date: '`r Sys.Date()`'
output:
  bookdown::word_document2: default
  bookdown::html_document2:
    number_sections: no
    toc: yes
    toc_float: yes
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: no
    toc: no
    keep_tex: yes
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

@@

We compiled functional traits from GLOPNET [@wright_worldwide_2004], TRY [@Kattge2011], and DRYAD [@chave_towards_2009] databases (see \@ref(tab:A2traits)). We kept dataset given by GLOPNET as origin dataset for observations. Dataset defined as origin corresponded to leaf lifespan ($LL$) and most of the time to leaf mass per area ($LMA$) and leaf nitrogen content per leaf dry mass ($Nmass$). We measured variable importance in functional traits to explain leaf lifespan with an out-of the bag method applied on a random forest. Then, we used a Bayesian approach to test different models with growing level of complexity. We retained the model with the best trade-off between model complexity (number of parameters $K$), convergence, likelihood, and prediction quality (root mean square error of prediction $RMSEP$). We finally tested the new allometry obtained with the selected model with TROLL simulations.

@@

Leaf lifespan $LL$ follows a log-normal law with location inferred from leaf lifespan $LMA$, nitrogen content $N$ and wood specific gravity $wsg$ and a scale $\sigma$. Each ${\beta_i}_d$ is following a normal law located on $\beta_i$ with a scale of $\sigma_i$. All $\beta_i$, $\sigma_i$, and $\sigma$ are assumed without preemption following a gamma law. $d$ represents the dataset random effects and encompass environmental and protocol variations.

@@

Simulations are validating that this new allometry resolve the issue of unrealistically low aboveground biomass for low LMA species due to an early death of individuals inside simulations. For instance with this allometry *Symphonia sp 1* (a low LMA species) is now reaching a realistic aboveground biomass above 400 $tonC.ha^{-1}$ and realistic diameter and age distribution inside the final population.
